---
title: "HipFractureEDA"
output: pdf_document
date: "2023-02-22"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(haven)
library(dplyr)
library(plyr)
library(ggplot2)
library(ggrepel)
library(corrplot)
library(patchwork)
library(psych)
library(tidyverse)
library(gridExtra)
library(naniar)
```

#Loading Data 
```{r cars}
# Each of us are saving our data locally (for now at least)
# To call out the sav file, we need to use different directories.

#hip_fracture_df = read_sav('../Analysis/Data/HipFractureQualityData.sav') # Bruno's Directory

hip_fracture_df = read_sav('/Users/zhixiangyao/Desktop/Spring\ 2023/Intro\ BioMed\ Data\ Sci/Project/Data/HipFractureQualityData.sav') # Carl's Directory
```


```{r cars}
#first off lets get a summary of the data to see what is missing. We want to look more closely at the encoded columns (Starting at Gender = 1) since we would not expect NA's here
options(max.print = 4000)
summary(hip_fracture_df)

#it looks like there are no NA's being displayed - it may be mis- encoding them when we read in the data

```

#Carl: It seems that the NA data are represented as -99



```{r}
# Deleting irrelavent data from the df
df <- hip_fracture_df[,!names(hip_fracture_df) %in% c("CASEID")]

# Keep an original df if further operations on the df will be applied.
df2 <- data.frame(df)
AGE_hist <- ggplot(df2, aes(x=AGE)) +
  geom_histogram(binwidth = 5) +
  labs(title = "Distribution of age", x = "Age", y = "Density") +
  theme(plot.title = element_text(hjust = 0.5))
ggsave("AGE_Histogram.png", plot = AGE_hist)
```

```{r}
df4 <- data.frame(df2)
df4[df4 == -99] <- NA
missing <- data.frame(map(df4, ~sum(is.na(.))))
missing[missing == 0] <- NA
missing <- missing[ , colSums(is.na(missing))==0]
```

```{r, echo = FALSE}
df2[sapply(df2, is.character)] <- lapply(df2[sapply(df2, is.character)], as.factor)
df2[sapply(df2, is.numeric)] <- lapply(df2[sapply(df2, is.numeric)], as.numeric)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Preprocessing the data by deleting the ID column

```{r}
head(hip_fracture_df,1) # Have a look at the names of each column
```



```{r}
RACE_df <- data.frame(table(df2['RACE_NEW']))
names(RACE_df)[names(RACE_df) == 'Var1'] <- 'Race'
RACE_df <- RACE_df %>%
  mutate(percentage = `Freq`/sum(`Freq`)) %>%
  arrange(percentage) %>%
  mutate(labels = scales::percent(percentage))
```

```{r}
Race_bar <- ggplot(RACE_df, aes(x = Race, y = percentage, fill = Race)) +
  scale_fill_brewer(palette = "Dark2") +
  geom_bar(stat="identity") +
  scale_y_continuous(labels=scales::percent) +
  labs(title="Race Distribution", y = "Percentage") +
  geom_text(aes(label = labels), nudge_y = 0.125, angle = -15) +
  theme(axis.text.x = element_text(angle = -30, hjust = 0.5))
ggsave("Race_Barchart.png", plot = Race_bar)
```

It maybe more reasonable to give a barchart showing certain complications among different race and sex, and age. For the purpose of our proposal, I will first pull up 4 figures.

CardiacComplications; InfectiousComplications

```{r}
# For CardiacComplications

agg_df_scr <- setNames(aggregate(df2$SEX, by=list(df2$SEX,df2$RACE_NEW,df2$CardiacComplications), FUN=length),
                   c("Sex", "Race","CardiacComplications","Counts"))

sex_cardiac_race <- ggplot(agg_df_scr, aes(x = Sex, fill = CardiacComplications))+
  geom_bar(mapping = aes(x = Sex, y = Counts, fill = CardiacComplications), stat = "identity", position = "fill")+
  facet_wrap(~Race, ncol = 3, nrow = 3, scales = 'free', labeller = label_wrap_gen(25)) +
  labs(title="Cardiac Complications of patients with differnt sex and race", x = "Sex", y = "Relative Proportion of Cardiac Complications") +
  theme(plot.title = element_text(hjust = 0.5), strip.text = element_text(size=7))
ggsave("Card_sex+race.png", plot = sex_cardiac_race)
```


```{r}
df3 <- data.frame(df2)

df3$Age_level <- cut(df3$AGE, breaks = c(50,75,85,100), labels=c("Age < 75","75 < Age < 85","Age > 85"))

agg_df_sca <- setNames(aggregate(df3$SEX, by=list(df3$SEX,df3$Age_level,df3$CardiacComplications), FUN=length),
                   c("Sex", "Age","CardiacComplications","Counts"))

sex_cardiac_age <- ggplot(agg_df_sca, aes(x = Sex, fill = CardiacComplications))+
  geom_bar(mapping = aes(x = Sex, y = Counts, fill = CardiacComplications), stat = "identity", position = "fill")+
  facet_wrap(~Age, ncol = 3, nrow = 1, scales = 'free') +
  labs(title="Cardiac Complications of patients with differnt sex and age", x = "Sex", y = "Relative Proportion of Cardiac Complications") +
  theme(plot.title = element_text(hjust = 0.5), strip.text = element_text(size=7))
ggsave("Card_sex+age.png", plot = sex_cardiac_age)
```

```{r}
# For CardiacComplications

agg_df_sir <- setNames(aggregate(df2$SEX, by=list(df2$SEX,df2$RACE_NEW,df2$InfectiousComplications), FUN=length),
                   c("Sex", "Race","InfectiousComplications","Counts"))

sex_infect_race <- ggplot(agg_df_sir, aes(x = Sex, fill = InfectiousComplications))+
  geom_bar(mapping = aes(x = Sex, y = Counts, fill = InfectiousComplications), stat = "identity", position = "fill")+
  facet_wrap(~Race, ncol = 3, nrow = 3, scales = 'free', labeller = label_wrap_gen(25)) +
  labs(title="Infectious Complications of patients with differnt sex and race", x = "Sex", y = "Relative Proportion of Infectious Complications") +
  theme(plot.title = element_text(hjust = 0.5), strip.text = element_text(size=7))
ggsave("infect_sex+race.png", plot = sex_infect_race)
```


```{r}
agg_df_sia <- setNames(aggregate(df3$SEX, by=list(df3$SEX,df3$Age_level,df3$InfectiousComplications), FUN=length),
                   c("Sex", "Age","InfectiousComplications","Counts"))

sex_infect_age <- ggplot(agg_df_sia, aes(x = Sex, fill = InfectiousComplications))+
  geom_bar(mapping = aes(x = Sex, y = Counts, fill = InfectiousComplications), stat = "identity", position = "fill")+
  facet_wrap(~Age, ncol = 3, nrow = 1, scales = 'free') +
  labs(title="Infectious Complications of patients with differnt sex and age", x = "Sex", y = "Relative Proportion of Infectious Complications") +
  theme(plot.title = element_text(hjust = 0.5), strip.text = element_text(size=7))
ggsave("infect_sex+age.png", plot = sex_infect_age)
```



```{r}
agg_df_srad <- setNames(aggregate(df2$SEX, by=list(df2$SEX,df2$RACE_NEW,df2$READMISSION1), FUN=length),
                   c("Sex", "Race","Readmission","Counts"))

sex_race_read <- ggplot(agg_df_srad, aes(x = Sex, fill = Readmission))+
  geom_bar(mapping = aes(x = Sex, y = Counts, fill = Readmission), stat = "identity", position = "fill")+
  facet_wrap(~Race, ncol = 3, nrow = 3, scales = 'free') +
  labs(title="Infectious Complications of patients with differnt sex and age", x = "Sex", y = "Relative Proportion of Infectious Complications") +
  theme(plot.title = element_text(hjust = 0.5), strip.text = element_text(size=7))
ggsave("readmit_sex+race.png", plot = sex_race_read)
```


















